---
title: "uwa6xv_M01_HW"
author: "Alanna Hazlett"
date: "2024-02-01"
output:
  pdf_document: default
  html_document: default
---
# Problem 1
**(a)**
**Create latest**
Read in UScovid.csv.
```{r}
USCovid.df<-read.csv("UScovid.csv",header=TRUE)
```
Get subset of data from June 3, 2021 and remove date and flip columns.
```{r}
USCovid1.df<-USCovid.df[c(1381437:1384683),c(2,3,5,6)]
```
Remove rows with Unknown counties. 
```{r}
UnknownCounty<-which(USCovid1.df$county=="Unknown")
USCovid2.df<-USCovid1.df[-c(91,181,316,321,386,535,551,692,884,1177,1199,1224,1241,1322,1405,1757,1791,1812,2054,2141,2222,2402,2414,2615,2902,2919),]
```
Order by county then state alphabetically. Name 'latest' and head() first 6 rows.
```{r}
latest<-USCovid2.df[order(USCovid2.df$county,USCovid2.df$state),]
head(latest)
```
**(b)**
County case fatality rate (numeric variable to numeric variable).
```{r}
death.rate<-(latest$deaths / latest$cases) * 100
death.rate<-round(death.rate,2)
```
Add death.rate to latest and display first 6 rows.
```{r}
latest<-data.frame(latest,death.rate)
head(latest)
```
**(c)**
Display the 10 largest cases by county.
```{r}
LargeCases<-latest[order(-latest$cases),]
head(LargeCases,10)
```
**(d)**
Display the 10 largest deaths by county.
```{r}
LargeDeaths<-latest[order(-latest$deaths),]
head(LargeDeaths,10)
```
**(e)**
Display the 10 largest case fatality rates by county.
```{r}
LargeDeathRate<-latest[order(-latest$death.rate),]
head(LargeDeathRate,10)
```
**(f)**
Display the counties with the 10 highest case fatality rates among counties with at least 100,000 cases.
```{r}
LargeCountyLargeRate<-latest[which(latest$cases>99999),]
LargeCountyLargeRate<-LargeCountyLargeRate[order(-LargeCountyLargeRate$death.rate),]
head(LargeCountyLargeRate,10)
```
**(g)**
Display the number of cases, deaths, and case fatality rates for Albemarle, Virginia and Charlottesville city, Virginia.
```{r}
Albemarle<-latest[c(35),]
head(Albemarle)
Charlottesville<-latest[c(474),]
head(Charlottesville)
```
# Problem 2 
**(a)**
**Create state.level** 
Select rows with June 3, 2021 date; select columns state, number of cases, and number of deaths.
```{r}
USCovid3.df<-USCovid.df[c(1381437:1384683),c(3,5,6)]
```
Create state.level with sums of cases and sums of deaths by state.
*Received help from Karunya Iyappan on Piazza to add state names column.* 
```{r}
state.names<-unique(USCovid3.df$state)
state.cases<-tapply(USCovid3.df$cases,USCovid3.df$state,sum)
state.deaths<-tapply(USCovid3.df$deaths,USCovid3.df$state,sum)
state.level<-data.frame(state.names,state.cases,state.deaths)
```
Remove row labels (non-working column). 
*Received help from Karunya Iyappan on Piazza to remove state row names.*
```{r}
rownames(state.level)<-c()
```
Renaming columns of state.level. 
```{r}
names(state.level)[c(1,2,3)]<-c("state","cases","deaths")
head(state.level)
```
**(b)**
State case fatality rate (numeric variable to numeric variable).
```{r}
state.rate<-(state.level$deaths / state.level$cases) * 100
state.rate<-round(state.rate,2)
```
Add state.rate to state.level and display the first 6 rows.
```{r}
state.level<-data.frame(state.level,state.rate)
head(state.level)
```
**(c)**
The case fatality rate for Virginia is 1.66%.
```{r}
Virginia<-state.level[c(51),]
head(Virginia)
```
**(d)**
The case fatality rate for Puerto Rico is N/A. The fatality rate can not be calculated, because there are no recorded deaths for Puerto Rico.
```{r}
Puerto_Rico<-state.level[c(42),]
head(Puerto_Rico)
```
**(e)**
Display the 10 largest state case fatality rates. The 10 states with the highest fatality rates (from highest to lowest) are New Jersey, Massachusetts, New York, Connecticut, District of Columbia, Mississippi, Pennsylvania, Lousiana, New Mexico, and Maryland.
```{r}
LargeStateRate<-state.level[order(-state.level$state.rate),]
head(LargeStateRate,10)
```
**(f)**
Display the 10 lowest state case fatality rates. The 10 states with the lowest case fatality rates (from lowest to highest) are Alaska, Utah, Virgin Islands, Vermont, Nebraska, Idaho, Northern Mariana Islands, Wisconsin, Wyoming, and Colorado.
```{r}
SmallStateRate<-state.level[order(state.level$state.rate),]
head(SmallStateRate,10)
```
**(g)**
Create csv called stateCovid.csv.
```{r}
write.csv(state.level,file="stateCovid.csv",row.names=TRUE)
```

