---
title: "M10Guided"
author: "Alanna Hazlett"
date: "2024-04-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(leaps)
library(car)
Data<-read.table("nfl.txt", header=TRUE, sep="")
```

# Problem 1

Create diagnostic plots for this regression. What are the plots telling us?
```{r}
result<-lm(y~x2+x7+x8,data=Data)
par(mfrow=c(2,2))
plot(result)
acf(result$residuals, main="ACF Plot of Residuals")
```
\
Assumption 1: Do the errors have mean of 0 for each value of the predictor -Yes\
Assumption 2: Do the errors have constant variance for each value of the predictor -Yes\
Assumption 3: Are the errors independent (acf plot) - Yes\
Assumption 4: Are the errors normally distributed? -Yes\
There is a linear relationship between the predictors x2 (Passing yards-Season), x7 (Percent rushing), x8 (Opponent's rushing yards) with our response variable (Games won).\

# Problem 2

Generate partial regression plots for each of the predictors. Interpret what these plots are telling us.
```{r}
car::avPlots(result)
```
\
UL: Linear pattern present, positive linear association. Slope of 250 = estimated coefficient of x2 in the model.\
UR: Linear pattern present, positive linear association. Slope of 0.2 = estimated coefficient of x7 in the model.\
LL: Linear pattern present, negative linear association. Slope of -0.005 = estimated coefficient of x8 in the model.\

# Problem 3

Using externally studentized residuals, do we have any outliers? What teams are these?\
```{r}
sort(abs(rstandard(result)))
```
*Guideline externally studentized residuals greater than 2 are flagged as outliers*\
Team 1 is a potential outlier.\

# Problem 4

Do we have any high leverage data points for this multiple linear regression? What teams are these?
```{r}
sort(lm.influence(result)$hat)
hii<-lm.influence(result)$hat
n<-nrow(Data)
p<-4
hii[hii>2*p/n]
```
*Guideline for significant leverage is hii > 2p/n = 2(4)/28 = \
My results show that there are no high leverage data points in this data set.\

# Problem 5

Use DFFITSi, DFBETASj,i, and Cook’s distance to check for influential observations. What teams are influential?\
**(DFFITS)**\
```{r}
p<-4
n<-nrow(Data)
DFFITS<-dffits(result)
DFFITS[abs(DFFITS)>2*sqrt(p/n)]
```
According to DFFITS we have no influential observations.\
**(DFBETAS)**\
```{r}
DFBETAS<-dfbetas(result)
abs(DFBETAS)>2/sqrt(n)
DFBETAS[10,]
DFBETAS[21,]
```
*Guideline of significance = 2/sqrt(n) = 0.3780*\
Team 10 coefficient for x8 = -0.4260\
Team 21 coefficient for x7 = -0.4148\
**(Cook's Distance)**\
```{r}
COOKS<-cooks.distance(result)
sort(COOKS)
#COOKS[COOKS>1]
```
According to cook's distance we have no influential observations (teams). 







