---
title: "M04Guided"
author: "Alanna Hazlett"
date: "2024-02-17"
output: html_document
---
```{r include=FALSE}
library(tidyverse)
library(palmerpenguins)
Data<-penguins
Gentoo<-Data %>% 
  filter(species=="Gentoo") %>% 
  filter(!is.na(flipper_length_mm)) %>% 
  filter(!is.na(body_mass_g))
```
# Problem 1

Produce a plot of body mass against flipper length for Gentoo penguins. Write the
estimated linear regression equation.
```{r}
result<-lm(body_mass_g~flipper_length_mm,data=Gentoo)
summary(result)
ggplot2::ggplot(Gentoo,aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  labs(x="Flipper Length (mm)",y="Body Mass (g)",title="Gentoo Species Body Mass against Flipper Length")
```
The estimated linear regression equation is:
$$\hat y=\hat\beta_{0}+\hat\beta_{1}x$$
$$\hat y = -6787.281 + 54.623x$$ 

# Problem 2

What is the change in the predicted body mass (in grams) when flipper length increases
by 1mm, for Gentoo penguins? Also report the corresponding 95% confidence interval
for the change in the predicted body mass (in grams) when flipper length increases by
1mm.\
\ 
The predicted body mass increases by 54.623 grams when flipper length increases by 1mm.
```{r}
confint(result,level=0.95)[2,]
```
 
# Problem 3  
 
Conduct a hypothesis test to determine whether or not there is a linear association
between body mass and flipper length for Gentoo penguins. State the hypotheses,
p-value, and conclusion in context.\

$$ H_{0}: \beta_{1} = 0\\H_{a}: \beta_{1} \neq 0 $$ 
\
The p-value is 2e-16, which is smaller than our significance level of 0.05. This means that we reject our null hypothesis of slope = 0 and our data supports the alternative hypothesis that slope does not equal zero. This indicates that there is a linear association between our variables. 
\

# Problem 4

Are your results from parts 2 and 3 consistent? Briefly explain.\
Yes, our hypothesis test indicates that slope does not equal zero.\ 
Our confidence interval shows that we have 95% confidence that our slope is within 44.67 and 64.58, and zero is not within this interval, so slope is not equal to zero. 

# Problem 5 

Estimate the mean body mass (in grams) for Gentoo penguins with flipper lengths of
200mm. Also report the 95% confidence interval for the mean body mass (in grams)
for Gentoo penguins with flipper lengths of 200mm
\
$$\hat y = -6787.281 + 54.623(200)\\
\hat y = 4137.319g$$ 
```{r}
newdata<-data.frame(flipper_length_mm = 200)
predict(result,newdata,level=0.95,interval="confidence")
```
\
The confidence interval is from 3954.45g to 4319.99g.\

# Problem 6

Report the 95% prediction interval for the body mass (in grams) of a Gentoo penguin
with flipper length of 200mm
```{r}
predict(result,newdata,level=0.95,interval="prediction")
```
\
The prediction interval is from 3401.12g to 4873.32g.\

# Problem 7

A researcher hypothesizes that for Gentoo penguins, the predicted body mass increases
by more than 50 g for each additional mm in flipper length. Conduct an appropriate
hypothesis test. What is the null and alternative hypotheses, test statistic, and conclusion?\
$$ H_{0}: \beta_{1} = 50\\H_{a}: \beta_{1} > 50 $$ 
The test statistic is:
$$ t = \frac {\hat \beta_{1} - 50} {se(\hat \beta_{1})}$$
$$ t =  \frac {54.623 - 50}{5.028} = 0.91945$$

```{r}
qt(0.95,121)
1-pt(0.9194511,df=123-2)
```
Our test statistic is less than our critical value, so we fail to reject the null hypothesis.\
The p-value is larger than our significance level of 0.05, so we fail to reject the null hypothesis.






